Modeling costs and benefits of genomic wastewater surveillance

### 2023-05-11 paper/model-documentation.qmd
updates model documentation

---
title: Modeling costs and benefits of genomic wastewater surveillance
subtitle: Supplementary appendix draft
#author:
#  - name: Pedro Nascimento de Lima
#    email: plima@rand.org
#    affiliations: 
#        - id: rand
#          name: RAND Corporation
#          department: Engineering and Applied Sciences
#          address: 1776 Main Street
#          city: Santa Monica
#          state: CA
#          postal-code: 90401
#    attributes:
#        corresponding: true
#  - name: and colleagues
#    email: others@rand.org
#    affiliations:
#        - ref: rand
abstract: |
  Introduces a model to assess the value provided by wastewater genomic surveillance network. This document is meant to be used as a supplementary appendix of the paper.
keywords: 
  - Genomic Surveillance
  - Pandemic Response
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  elsevier-html:
    keep-tex: true
    mermaid:
      theme: neutral
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  docx:
      cite-style: super
  html:
    keep-tex: true
    mermaid:
      theme: neutral
---

@@

@@

## Introducing and removing NPIs

@@

>>>>

---
title: Modeling costs and benefits of genomic wastewater surveillance
# subtitle: Supplementary appendix draft
#author:
#  - name: Pedro Nascimento de Lima
#    email: plima@rand.org
#    affiliations: 
#        - id: rand
#          name: RAND Corporation
#          department: Engineering and Applied Sciences
#          address: 1776 Main Street
#          city: Santa Monica
#          state: CA
#          postal-code: 90401
#    attributes:
#        corresponding: true
#  - name: and colleagues
#    email: others@rand.org
#    affiliations:
#        - ref: rand
abstract: |
  Introduces a model to assess the value provided by wastewater genomic surveillance network. This document is meant to be used as a supplementary appendix of the paper.
keywords: 
  - Genomic Surveillance
  - Pandemic Response
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  elsevier-html:
    keep-tex: true
    mermaid:
      theme: neutral
    journal:
      #name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  docx:
      cite-style: super
  html:
    keep-tex: true
    mermaid:
      theme: neutral
---

@@

# Background

While researchers have demonstrated that wastewater-based surveillance (WBS) systems can provide an early warning system for pandemic response [@zhu2021] and be used to support estimation of the latent number of infections in COVID-19 transmission models [@soller2022], the value of information provided by WBS has yet to be demonstrated.

This paper develops a stochastic metapopulation disease transmission model and uses it to assess the conditions under which WBS systems can provide value to local public health decision makers who respond to a pandemic using surveillance data. We evaluate a range of scenarios that represent different potential pathogens and highlight scenarios under which WBS systems may or may not be cost-effective.

@@

## Controlling NPIs

@@

## Cost of interventions

The cost of reducing infectious contacts over a time period $t_0$ to $t_f$ may be defined as a function of the effect of interventions

$$
C = \int_{t_0}^{t_f}{f(\mathbf{l}_t\tau)}
$${#eq-cost}

where $f(x)$ is a function that maps a percent reduction in potentially infectious contacts to a monetary value. At the time of the intervention, decision-makers do not have precise information about the costs of interventions $\tau$ or its costs represented by function $f(x)$.
